---
title: "AAMC Table A-1 (MD Schools) and Table B-8 (MD/PhD Schools) Visualizer"
author: "Author: toofastdan"
date: "Date Modified: 1/16/2022"
data source: "ijhua on GitHub: https://github.com/ijhua/aamc_admissions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(readxl)
library(writexl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
```

```{r include=FALSE}
# Setting the file directory with the excel files created by the python table analyzer
main_dir <- "C:/Users/TooFastDan/Documents/MD_PhD Application/Python Analysis/AAMC data"

# Importing MD table A-1 information
md <- read_xlsx(paste0(main_dir, "/AAMC Table A-1 Merged.xlsx"))
md$degree <- "MD"
md_totals <- read_xlsx(paste0(main_dir, "/AAMC Table A-1 Totals.xlsx"))
md_totals$degree <- "MD"

# Importing MD/PhD table B-8 information
mdphd <- read_xlsx(paste0(main_dir, "/AAMC Table B-8 Merged.xlsx"))
mdphd$degree <- "MD-PhD"
mdphd_totals <- read_xlsx(paste0(main_dir, "/AAMC Table B-8 Totals.xlsx"))
mdphd_totals$degree <- "MD-PhD"

# Merging dataframes to make things easier to graph
a1b8 <- rbind(md, mdphd) %>% as.data.frame()
a1b8$degree <- factor(x = a1b8$degree, levels = c("MD", "MD-PhD"))
a1b8_totals <- rbind(md_totals, mdphd_totals) %>% as.data.frame()
a1b8_totals$degree <- factor(x = a1b8_totals$degree, levels = c("MD", "MD-PhD"))
```


## Graphing the Number of Total Applications/Applicants/Matriculants

```{r include=FALSE}
a1b8_totals_apps <- a1b8_totals[ , c(1, 11, 15)]
a1b8_totals_apps$metric <- "Total Applications"
colnames(a1b8_totals_apps) <- c("n", "year", "degree", "metric")

a1b8_totals_applicants <- a1b8_totals[ , c(12, 11, 15)]
a1b8_totals_applicants$metric <- "Applicants"
colnames(a1b8_totals_applicants) <- c("n", "year", "degree", "metric")

a1b8_totals_matriculants <- a1b8_totals[ , c(6, 11, 15)]
a1b8_totals_matriculants$metric <- "Matriculants"
colnames(a1b8_totals_matriculants) <- c("n", "year", "degree", "metric")

a1b8_totals_bulk <- rbind(a1b8_totals_apps, a1b8_totals_applicants, a1b8_totals_matriculants)
a1b8_totals_bulk$degree_metric <- paste0(a1b8_totals_bulk$degree, ": ", a1b8_totals_bulk$metric)
a1b8_totals_bulk$degree_metric <- factor(x = a1b8_totals_bulk$degree_metric, levels = c("MD: Total Applications", "MD: Applicants", "MD: Matriculants", "MD-PhD: Total Applications", "MD-PhD: Applicants", "MD-PhD: Matriculants"))
```

```{r include=FALSE}
layout_ggplotly <- function(ggplotly_graph, x, y){
  # Changing the x and y axes positions in a ggplotly graph
  # The 1 and 2 goes into the list that contains the options for the x and y axis labels respectively
  ggplotly_graph[['x']][['layout']][['annotations']][[1]][['y']] <- y
  ggplotly_graph[['x']][['layout']][['annotations']][[2]][['x']] <- x
  return(ggplotly_graph)
}
```

```{r echo=FALSE}
a1b8_totals_bulk_graph <- a1b8_totals_bulk %>% ggplot(aes(x = year, y = n)) +
  geom_line() +
  geom_point() + 
  facet_wrap(~degree_metric, nrow = 2, ncol = 3, scales = "free_y") +
  labs(title = "MD and MD/PhD Applicants and Matriculants Over Time", x = "Year", y = "Number") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
#print(a1b8_totals_bulk_graph)
gp <- ggplotly(a1b8_totals_bulk_graph, width = 1200, height = 700)
gp <- layout_ggplotly(ggplotly_graph = gp, x = -0.055, y = -0.04)
gp
```


## Graphing the Applicants & Matriculants Together

```{r echo=FALSE}
a1b8_totals_bulk2 <- a1b8_totals_bulk[a1b8_totals_bulk$metric!="Total Applications", ]
a1b8_totals_bulk2$metric <- factor(x = a1b8_totals_bulk2$metric, levels = c("Applicants", "Matriculants"))

a1b8_totals_bulk2_graph <- a1b8_totals_bulk2 %>% ggplot(aes(x = year, y = n, color = metric)) +
  geom_line() +
  geom_point() + 
  facet_wrap(~degree, nrow = 1, ncol = 2, scales = "free_y") +
  scale_color_manual(values = c("#ff4747", "#7f6ce0")) +
  labs(title = "MD and MD/PhD Applicants and Matriculants Over Time", x = "Year", y = "Number of Applicants/Matriculants") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
#print(a1b8_totals_bulk2_graph)
gp <- ggplotly(a1b8_totals_bulk2_graph, width = 1200, height = 550)
gp <- layout_ggplotly(ggplotly_graph = gp, x = -0.05, y = -0.04)
gp
```

MD:  There was a `r round((a1b8_totals[10, 12] - a1b8_totals[9, 12]) / (a1b8_totals[9, 12]) * 100, digits = 1)`% increase in the 2021 cycle, relative to the 2020 cycle.

MD/PhD:  There was a `r round((a1b8_totals[20, 12] - a1b8_totals[19, 12]) / (a1b8_totals[19, 12]) * 100, digits = 1)`% increase in the 2021 cycle, relative to the 2020 cycle.


## Graphing in-state and out-of-state percentages over time

```{r echo=FALSE}
a1b8_totals_state <- tidyr::pivot_longer(data = a1b8_totals, cols = c("in state applicants", "in state matriculants"), names_to = "Application_Status", values_to = "percent_instate")
a1b8_totals_state$Application_Status <- gsub(pattern = "in state ", replacement = "", x = a1b8_totals_state$Application_Status)
a1b8_totals_state$Application_Status <- factor(x = a1b8_totals_state$Application_Status, levels = c("applicants", "matriculants"))

a1b8_totals_state_graph <- a1b8_totals_state %>% ggplot(aes(x = cycle_year, y = percent_instate, color = Application_Status)) +
  geom_line() +
  geom_point() + 
  ylim(0, 65) +
  scale_color_manual(values = c("#ff4747", "#7f6ce0")) +
  facet_wrap(~degree, nrow = 1, ncol = 2, scales = "fixed") +
  labs(title = "MD and MD/PhD Percent in-state Over Time", x = "Year", y = "Percent in-state Applicants/Matriculants") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
#print(a1b8_totals_state_graph)
gp <- ggplotly(a1b8_totals_state_graph, width = 1200, height = 550)
gp <- layout_ggplotly(ggplotly_graph = gp, x = -0.025, y = -0.04)
gp
```

## Graphing percent sex percentages over time

```{r echo=FALSE}
a1b8_totals_sex <- tidyr::pivot_longer(data = a1b8_totals, cols = c("women applicants", "men applicants", "women matriculants", "men matriculants"), names_to = "Application_Status", values_to = "percent_sex")
a1b8_totals_sex$Application_Status <- factor(x = a1b8_totals_sex$Application_Status, levels = c("women applicants", "women matriculants", "men applicants", "men matriculants"))

a1b8_totals_sex_graph <- a1b8_totals_sex %>% ggplot(aes(x = cycle_year, y = percent_sex, color = Application_Status)) +
  geom_line() +
  geom_point() + 
  #ylim(0, 65) +
  scale_color_manual(values = c("#fbadff", "#ff00ee", "#adc8ff", "#000bff")) +
  facet_wrap(~degree, nrow = 1, ncol = 2, scales = "fixed") +
  labs(title = "MD and MD/PhD Percent Sex Over Time", x = "Year", y = "Percent Sex Applicants/Matriculants") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
#print(a1b8_totals_sex_graph)
#ggsave(plot = a1b8_totals_sex_graph, filename = paste0(main_dir, "/percent sex over time.png"), width = 3000, height = 1450, units = "px", dpi = 300)
gp <- ggplotly(a1b8_totals_sex_graph, width = 1200, height = 550)
gp <- layout_ggplotly(ggplotly_graph = gp, x = -0.025, y = -0.04)
gp
```

## Visalizing my cycle with a line graph

```{r}

```

























